Rapid technological progress in white light-emitting diodes and its sources in innovation and technology spillovers

### 2023-05-29 2_SSL_NE/article/article_main.tex
updated main document

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% This is a (brief) model paper using the achemso class
%% The document class accepts keyval options, which should include
%% the target journal and optionally the manuscript type. 
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\documentclass[journal=jacsat,manuscript=article]{achemso}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% Place any additional packages needed here.  Only include packages
%% which are essential, to avoid problems later. Do NOT use any
%% packages which require e-TeX (for example etoolbox): the e-TeX
%% extensions are not currently available on the ACS conversion
%% servers.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\usepackage[version=3]{mhchem} % Formula subscripts using \ce{}
\usepackage{tabularx}
\usepackage[usenames,dvipsnames]{xcolor}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% If issues arise when submitting your manuscript, you may want to
%% un-comment the next line.  This provides information on the
%% version of every file you have used.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%\listfiles

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% Place any additional macros here.  Please use \newcommand* where
%% possible, and avoid layout-changing macros (which are not used
%% when typesetting).
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% Meta-data block
%% ---------------
%% Each author should be given as a separate \author command.
%%
%% Corresponding authors should have an e-mail given after the author
%% name as an \email command. Phone and fax numbers can be given
%% using \phone and \fax, respectively; this information is optional.
%%
%% The affiliation of authors is given after the authors; each
%% \affiliation command applies to all preceding authors not already
%% assigned an affiliation.
%%
%% The affiliation takes an option argument for the short name.  This
%% will typically be something like "University of Somewhere".
%%
%% The \altaffiliation macro should be used for new address, etc.
%% On the other hand, \alsoaffiliation is used on a per author basis
%% when authors are associated with multiple institutions.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\author{Michael Weinold}
\email{mweinold@ethz.ch}
\affiliation[PSI]{Technology Assessment Group, Paul Scherrer Institute, Villigen, Switzerland}
\alsoaffiliation[ETHZ]
{Chair for Entrepreneurial Risk, ETH Zurich, Zurich, Switzerland}
\alsoaffiliation[UCAM]
{Cambridge Centre for Environment, Energy and Natural Resource Governance, Department of Land Economy, University of Cambridge, Cambridge, United Kingdom}

\author{Sergey Kolesnikov}
\affiliation[UCAM]
{Cambridge Centre for Environment, Energy and Natural Resource Governance, Department of Land Economy, University of Cambridge, Cambridge, United Kingdom}

\author{Laura Diaz Anadon}
\affiliation[UCAM]{Cambridge Centre for Environment, Energy and Natural Resource Governance, Department of Land Economy, University of Cambridge, Cambridge, United Kingdom}
\alsoaffiliation[UHarv]
{Belfer Center for Science and International Affairs, Harvard University, Cambridge MA, USA}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% The document title should be given as usual. Some journals require
%% a running title from the author: this should be supplied as an
%% optional argument to \title.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\title[]{Rapid technological progress in white light-emitting diodes and its sources in innovation and technology spillovers}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% Some journals require a list of abbreviations or keywords to be
%% supplied. These should be set up here, and will be printed after
%% the title and author information, if needed.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\abbreviations{IR,NMR,UV}
\keywords{American Chemical Society, \LaTeX}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% The manuscript does not need to include \maketitle, which is
%% executed automatically.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{document}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% The "tocentry" environment can be used to create an entry for the
%% graphical table of contents. It is given here as some journals
%% require that it is printed as part of the abstract page. It will
%% be automatically moved as appropriate.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% The abstract environment will automatically gobble the contents
%% if an abstract is not used by the target journal.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{abstract}
  The first commercial white light-emitting diodes (LEDs) were introduced to the market in 1996. Since then, white LEDs have experienced major improvements in performance and efficiency. Significant decreases in cost have been driven by a rapid expansion of the market share of LED-based solid-state lighting (SSL), which in turn has driven down costs even further. Today, the technology is one of the key solutions in global climate change mitigation efforts. However, despite its significance, the extent and sources of the underlying improvements have not been systematically investigated. Understanding what has driven rapid progress in LEDs can provide lessons for accelerating innovation in a broad range of other demand-side clean energy technologies. With this aim, we gather and analyse systematic evidence on cost and performance improvements in white LEDs from cost and performance modelling, augmented by a literature review and interviews with experts in industry and academia. We find that the overall efficiency of the highest performing warm white LED packages has improved from $\eta_L=5.8\%$ in 2003 to $\eta_L=38.8\%$ in 2020. During the same period, we estimate that the cost of manufacturing low-power and mid-power LED packages at a U.S. location using state-of-the-art equipment has dropped from 1.1\$ to 0.05\$ (in 2020 USD) between 2003 and 2020, a 95.5\% decrease. We further show that technology spillovers — i.e., knowledge originating in other technologies, sectors, or scientific disciplines — affected all performance dimensions of LEDs, contributing no less than 8.5\% to total efficiency improvements and $\sim$100\% to the improvements in consumer experience metrics identified, thereby playing a key role in the widespread adoption of the technology. The increase in wafer size and manufacturing yield improvements were the primary causes of LED manufacturing cost reductions.
\end{abstract}

@@

\begin{table}[h!]
    \footnotesize
    \centering
    \begin{tabularx}{\textwidth}{|l|l|l|X|X|l|}
    \hline
        \textbf{Year} & \textbf{Desig.} & \textbf{Chemical formula} & \textbf{Description} & \textbf{Significance} & \textbf{SI Sec.} \\ \hline
        1996 & YAG & $Y_3 Al_5 O_{12}:Ce$ & Yttrium aluminium garnet (YAG) phosphor activated with cerium & First LED phosphor, enabled white LEDs & SI 3.2.2.1 \\ \hline
        1996 & YGAG & $(Y_{1-x} Gd_x)_3 Al_5 O_{12}:Ce$ & Gadolinium-doped YAG phosphor & First red-shifted phosphor, enabled warm white LEDs & SI 3.2.2.1 \\ \hline
        2002 & 258 & $(Ba,Sr)_2 Si_5 N_8:Eu^{2+}$ & Europium-doped nitridosilicate phosphor & First red LED phosphor & SI 3.2.2.2 \\ \hline
        2003 & QD & N/A & Quantum dot-based phosphor & First use of QD for LED light down conversion & SI 3.2.2.4 \\ \hline
        2005 & PFS & $K_2 SiF_6: Mn^{4+}$ & Manganese-activated potassium fluorosilicate (PFS) phosphor & First ultra-narrow-band red LED phosphor & SI 3.2.2.3 \\ \hline
        2013 & SLA & $Sr[Li Al]_3 N_4 ]:Eu^{2+}$ & Europium-doped cuboidal nitridolithoaluminate phosphor & Improved narrow-band red phosphor & SI 3.2.2.2 \\ \hline
        2016 & SALON & $Sr[Li_2 Al_2 O_2 N_2]:Eu^{2+}$ & Europium-doped oxonitride phosphor & High-performance ultra-narrow-band red phosphor & SI 3.2.2.2 \\ \hline
    \end{tabularx}
    \caption{LED down-conversion materials (phosphors and quantum dots) related to improvements in consumer experience metrics identified in this study. The \textit{Year} column represents the earliest reported application of invention in white LEDs. These differ from the years used in Figure \ref{fgr:consumer_experience} which correspond to the earliest publication of spectral data for representative LED products that relied on those materials. The \textit{SI Sec.} column refers to the corresponding section in the Supplementary Information that contains references and a detailed description of the history of corresponding invention. Abbreviations: Desig. - Designation, Sec. - Section.}
    \label{tab:phosphors}
\end{table}

@@

    \begin{tabularx}{\textwidth}{|l|l|l|X|X|X|l|X|}

@@

    \end{tabularx}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% The "Acknowledgement" section can be given in all manuscript
%% classes.  This should be given within the "acknowledgement"
%% environment, which will make the correct section or running title.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{acknowledgement}

@@

\end{acknowledgement}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% The same is true for Supporting Information, which should use the
%% suppinfo environment.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{suppinfo}

@@

\end{suppinfo}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% The appropriate \bibliography command should be placed here.
%% Notice that the class file automatically sets \bibliographystyle
%% and also names the section correctly.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\bibliography{bibliography}

>>>>

\documentclass[parskip=full]{article}

% functionality
\input{auxiliary/preamble/1_preamble_functionality_main}

% layout
\input{auxiliary/preamble/2_preamble_layout_report}

% bibliography
\input{auxiliary/preamble/1_preamble_functionality_bibliography}
\addbibresource{bibliography.bib}

%% METADATA %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\hypersetup{
    pdftitle={Hybrid Life-Cycle Methods},
    pdfauthor={Michael Weinold, Guillaume Majeau-Bettez, Russell McKenna},
}

%% MAIN DOCUMENT %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{document}

M. Weinold$^{1,2,3}$, S. Kolesnikov$^3$, L.D. Anadon$^{3,4}$ \\
        "Rapid technological progress in white light-emitting diodes and its sources in innovation and technology spillovers" \textit{Environmental Science \& Technology} (2023) \\
        \newline
        $^1$ Technology Assessment Group, Paul Scherrer Institute, Switzerland \\
        $^2$ ETH Zurich, Zurich, Switzerland \\
        $^3$ CEENRG, Dept. of Land Economy, University of Cambridge, UK \\
        $^4$ Belfer Center for Science and International Affairs, Harvard University, Cambridge MA, USA

\begin{abstract}
  The first commercial white light-emitting diodes (LEDs) were introduced to the market in 1996. Since then, white LEDs have experienced major improvements in performance and efficiency. Significant decreases in cost have been driven by a rapid expansion of the market share of LED-based solid-state lighting (SSL), which in turn has driven down costs even further. Today, the technology is one of the key solutions in global climate change mitigation efforts. However, despite its significance, the extent and sources of the underlying improvements have not been systematically investigated. Understanding what has driven rapid progress in LEDs can provide lessons for accelerating innovation in a broad range of other demand-side clean energy technologies. With this aim, we gather and analyse systematic evidence on cost and performance improvements in white LEDs from cost and performance modelling, augmented by a literature review and interviews with experts in industry and academia. We find that the overall efficiency of the highest performing warm white LED packages has improved from $\eta_L=5.8\%$ in 2003 to $\eta_L=38.8\%$ in 2020. During the same period, we estimate that the cost of manufacturing low-power and mid-power LED packages at a U.S. location using state-of-the-art equipment has dropped from 1.1\$ to 0.05\$ (in 2020 USD) between 2003 and 2020, a 95.5\% decrease. We further show that technology spillovers — i.e., knowledge originating in other technologies, sectors, or scientific disciplines — affected all performance dimensions of LEDs, contributing no less than 8.5\% to total efficiency improvements and $\sim$100\% to the improvements in consumer experience metrics identified, thereby playing a key role in the widespread adoption of the technology. The increase in wafer size and manufacturing yield improvements were the primary causes of LED manufacturing cost reductions.
\end{abstract}

@@

\begin{table}[h!]
    \footnotesize
    \centering
    \begin{NiceTabularX}{\textwidth}{|l|l|l|X|X|l|}
    \hline
        \textbf{Year} & \textbf{Desig.} & \textbf{Chemical formula} & \textbf{Description} & \textbf{Significance} & \textbf{SI Sec.} \\ \hline
        1996 & YAG & $Y_3 Al_5 O_{12}:Ce$ & Yttrium aluminium garnet (YAG) phosphor activated with cerium & First LED phosphor, enabled white LEDs & SI 3.2.2.1 \\ \hline
        1996 & YGAG & $(Y_{1-x} Gd_x)_3 Al_5 O_{12}:Ce$ & Gadolinium-doped YAG phosphor & First red-shifted phosphor, enabled warm white LEDs & SI 3.2.2.1 \\ \hline
        2002 & 258 & $(Ba,Sr)_2 Si_5 N_8:Eu^{2+}$ & Europium-doped nitridosilicate phosphor & First red LED phosphor & SI 3.2.2.2 \\ \hline
        2003 & QD & N/A & Quantum dot-based phosphor & First use of QD for LED light down conversion & SI 3.2.2.4 \\ \hline
        2005 & PFS & $K_2 SiF_6: Mn^{4+}$ & Manganese-activated potassium fluorosilicate (PFS) phosphor & First ultra-narrow-band red LED phosphor & SI 3.2.2.3 \\ \hline
        2013 & SLA & $Sr[Li Al]_3 N_4 ]:Eu^{2+}$ & Europium-doped cuboidal nitridolithoaluminate phosphor & Improved narrow-band red phosphor & SI 3.2.2.2 \\ \hline
        2016 & SALON & $Sr[Li_2 Al_2 O_2 N_2]:Eu^{2+}$ & Europium-doped oxonitride phosphor & High-performance ultra-narrow-band red phosphor & SI 3.2.2.2 \\ \hline
    \end{NiceTabularX}
    \caption{LED down-conversion materials (phosphors and quantum dots) related to improvements in consumer experience metrics identified in this study. The \textit{Year} column represents the earliest reported application of invention in white LEDs. These differ from the years used in Figure \ref{fgr:consumer_experience} which correspond to the earliest publication of spectral data for representative LED products that relied on those materials. The \textit{SI Sec.} column refers to the corresponding section in the Supplementary Information that contains references and a detailed description of the history of corresponding invention. Abbreviations: Desig. - Designation, Sec. - Section.}
    \label{tab:phosphors}
\end{table}

@@

    \begin{NiceTabularX}{\textwidth}{|l|l|l|X|X|X|l|X|}

@@

    \end{NiceTabularX}

@@

@@

@@

\newpage
\printbibliography
